### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd

---
title: "MS_Aims"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

In this paper, we <!--  MOVED_EDITED_START -->consider literature on modelling practice and draw on our prior experience with developing and using CHEMs to propose<!--  MOVED_EDITED_START -->:  (i) a set of ethical responsibilities for CHEM project teams<!-- and criteria for assessing ethical CHEM implementations -->;<!--  MOVED_EDITED_START --> (ii) attributes of CHEMs that can suggest fulfillment of these responsibilities; and (iii) criteria against which these attributes can be assessed.  <!--  MOVED_EDITED_END -->  
<!-- (i) some core ethical responsibilities of CHEM project teams;

 <!-- -->

<!-- (ii) describe a prototype software framework we have developed to support the ethical implementation of CHEMs; and -->

<!-- (iii)	assess an initial set of youth mental health CHEMs developed with our software framework against our proposed ethical CHEM implementation criteria.  -->