### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd

---
title: "MS_Discussion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

Ethical practice is a core expectation of health researchers and computational methods underpin most quantitative research, yet an understanding of what constitutes ethical public health modelling practice is underdeveloped [@10.3389/fpubh.2017.00068]. We have addressed this gap with our proposed responsibilities for model developers and funders, enabling CHEM attributes and CHEM implementation assessment criteria.

Currently, many if not most existing CHEMs are insufficiently transparent [@Jalali2021; @McManus2019; @Bermejo2017; @Ghabri2019], not reusable [@Feenstra2022; @Emerson2019], nor updatable [@Sampson2017; @10.3389/fpubh.2022.899874]. Addressing these issues will require changes to the tools, practices and funding underpinning health economic model development.

<!-- As illustrated by Table \@ref(tab:checktb), we have developed a prototype software framework that can be used to implement youth mental health CHEMs that largely satisfy our TRU criteria. However, in its current form our software framework is too fragile to be recommended as anything more than a prototype for supporting the development needs of other modelling teams and projects. We developed our prototype software framework primarily to address the user requirements of one group – ourselves. Factors such as user enjoyment, usability, active user-community and supporting resources influence adoption of software frameworks [@myllarniemi2018development]. We currently lack the resources required to  address these issues for our software framework. -->

<!-- However, our prototype has a number of features that future work on CHEM software frameworks may find useful to incorporate. Firstly, developing a software framework to work within an existing and widely used open source programming language such as R, can keep framework scope relatively narrow. This makes it more tractable to develop, maintain and learn, and facilitates integration with other modelling tools written in that language (e.g., the dependency libraries we list in Table \@ref(tab:cpkgs)). Secondly, implementation that combines both object-oriented and functional programming paradigms can avail of the modular and syntactical simplicity benefits of the former, while limiting needless bundling of code artefacts. Thirdly, a sensible trade-off needs to be found between transparent code implementation (which requires clear and sufficiently detailed documentation) and Agile Software Development (for which a foundational principle is prioritising the development of working code over writing documentation [@beck2001manifesto]). Our software framework makes this trade off by enforcing the use of consistent code style conventions and file organisation which in turn enables automated generation of simple documentation at every CHEM module update. All model data-structures and algorithms are therefore always documented (at least minimally, with machine authored content), meaning model developers have a requirement to write customised documentation less frequently. Many of the features of our prototype software framework have the potential to be significantly enhanced through integration with evolving code and documentation authoring capabilities of large language models. -->

<!-- Our software framework provides the outline of a technical solution to some of the challenges of open source CHEM development. However, better tools alone are insufficient and policy and institutional changes are also required.  -->
Reducing waste in research is a core responsibility of research funders [@chalmers2014increase] and funding the development of CHEMs that are not adequately understood, reused or updated is wasteful. Existing incentive structures for health economists generally do not promote facilitating peers to reuse their work and funding is rarely adequate to maintain and routinely update a model. Currently, it can take “an extraordinary amount of idealism” to commit to authoring and maintaining research software [@anzt2020environment]. Previously recommended strategies for more beneficial health economic research investments include support for harmonised ethical standards for model development [@10.3389/fpubh.2017.00068], methodological innovation to improve model transferability [@craig2018taking], networks of modellers working on common health conditions [@Sampson2019], and  centralized infrastructure such as open source model repositories [@Pouwels2022] and a standard platform for model implementations [@Ghabri2019]. Development of software frameworks to support ethical CHEM implementations could enable and enhance each of these strategies.

<!-- A future software framework for ethical CHEMs would ideally incorporate a base set of features useful to developers of computational models across all domains of public health and health economics, with the capability for community-led extensions that are tailored to the needs of modellers focused on specific health-conditions. Such approaches will likely be important to ensure a sufficient potential user-base to make CHEM software frameworks sustainable. -->