### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Imp_Example.Rmd

---
title: "MS_Imp_Guidelines"
output: html_document
date: '2022-07-21'
---

We are using the ready4 software framework to implement multiple transparent, reusable and updatable CHEMs in youth mental health (see Figure \@ref(fig:fig1)).  Some initial outputs from this work are publicly available as development version releases.

```{r, child = child_figs_ls$fig1, echo=FALSE}
```

##	Economic topics
Currently, we are using the ready4 software framework to develop, apply, and share youth mental health CHEMs in four of the twelve domains of health economics identified by Wagstaff and Culyer [@wagstaff2012four]:

-	health and its value (our projects: models to map psychological and functional measures to health utility);

-	determinants of health and ill-health (our projects: models for creating synthetic household populations with key risk and protective factors for mental disorders);

-	demand for health and health care (our projects: spatial epidemiology and help-seeking choice models); and

-	supply of health services (our projects: a model of primary mental health care services for young people).

Potential future directions are to supplement this work with CHEMs in two additional Wagstaff and Culyer domains of: (i) public health (to model the impact of selected fiscal policy and regulation options on young people's mental health); and (ii) human resources (to model the supply and behaviours of the youth mental health workforce). Our ultimate aim is to flexibly combining all our CHEMs in analyses that help answer questions in two additional Wagstaff and Culyer domains:

-	efficiency and equity (our goal: assess the distributional impacts and identify the optimal targeting of youth mental health interventions); and

-	economic evaluation (our goal: assess the cost-utility of competing policy options for improving the mental health of young people).

Although we are principally interested in using CHEMs to answer policy questions relating to the mental health of young people in Australia, we want to facilitate CHEM transferability to other jurisdictions. Our CHEMs are being derived from and applied to real data (which can be empirical, simulated or assumption, so long as it is appropriate for use in analysis intended to inform decision-making) from Australia. To help demonstrate the potential use of CHEMs in other decision contexts, we also create toy datasets. Data created for illustration purposes is prominently labelled as not for use in decision-analysis.

## Case study: health and its value
We have previously described a study [@Hamilton2021.07.07.21260129] to develop utility mapping models for use in samples of young people presenting to primary mental health services. The ready4 software framework was used in that study to develop CHEM modules, supply those modules with data and implement modelling analyses, creating the following artefacts:

 - development version module libraries for describing and validating youth mental health human record datasets [@hamilton_matthew_2022_6084467], scoring health utility [@hamilton_matthew_2022_6084824], specifying utility mapping models [@hamilton_matthew_2022_6116701] and implementing reproducible utility mapping studies [@gao_caroline_2022_6130155];

 - a development version library of functions for finding and using utility mapping models developed with these tools [@matthew_p_hamilton_2021_5646669];

 - collections of real data (study input and results [@DVN/DKDIB0_2021]) and toy data (synthetic populations for testing model modules [@DVN/HJXYKQ_2021]);

 - programs for replicating all steps from data ingest to manuscript production [@hamilton_matthew_2022_6129906], applying utility mapping models to new data [@hamilton_matthew_2022_6416330] and generating a synthetic representation of the study dataset [@hamilton_matthew_p_2022_6321821]; and

 - subroutines for creating a catalogue of utility mapping models [@hamilton_matthew_2022_6116385] and generating a draft scientific manuscript [@matthew_p_hamilton_2022_5976988] for studies implemented with these modules.

We created a checklist (Table \@ref(tab:checktb)) that we used to subjectively assess these study outputs against TRU criteria. For each criterion, we provided a global assessment of whether it was met using the responses “yes”, “no” or “partial”.  We believe our utility mapping study CHEMs have satisfactorily met five of the six criteria (T1, T2, R1, R2 and U1) and have partially met one criterion (U2). The main shortcomings that we identified when applying the assessment criteria was that we have yet to adequately implement unit testing of the R libraries authored as part of this study.

\blandscape

```{r, child = child_docs_ls$four, echo=FALSE}
```

\elandscape