### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd

---
title: "MS_Imp_Guidelines"
output: html_document
date: '2022-07-21'
---

To support the development of CHEMs that meet TRU assessment criteria, we have created a prototype software framework called ready4. The software framework aims to provide a toolkit for: (i) enabling modular implementation of CHEMs; (ii) authoring and documenting CHEM modules; (iii) managing the labelling and transfer of CHEM input and output data; and (iv) authoring reproducible analyses that apply CHEMs to compatible data.

To achieve these goals, we have implemented the software framework as R [@RCORE2022] code libraries that integrate with a number of online services and which are supported by a documentation website.

## R libraries 
A library in the R language will typically depend on multiple other R libraries, all of which potentially have different authors. As the number of third-party dependencies of an R library grows so does the fragility of that library (e.g., the library may cease to work as intended due to changes in one of its dependency libraries). To reduce the fragility of our framework we implemented it as multiple R libraries rather than one R library. In total we authored six novel R libraries to implement the ready4 framework, all of which have distinct purposes and dependencies (Table \@ref(tab:cpkgs)).

One framework library provides a *foundation for modular CHEM implementations*. In modular model implementations, modules need to be able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering]. To achieve this, the foundation framework library defines a template CHEM module (using R’s S4 class system), which can be used to create other CHEM modules with a common set of inherited properties. One of these inherited properties is a novel syntax of 15 core commands that enable CHEM module algorithms to be consistently named. The foundation library also contains tools for retrieving web-based information on CHEM modules, datasets and analysis programs authored with the framework and tools for partially automating updates to a project documentation website.

Three framework libraries are designed to help streamline and standardise workflows for *authoring CHEM modules* from the template module. The R language supports functional and object-oriented programming paradigms [@10.1214/13-STS452]. Authoring with each of these paradigms is facilitated by a dedicated CHEM module authoring library. One module authoring library contains tools for writing functions in a consistent house style and automatically generating basic documentation for each function. A second module authoring library contains tools to help streamline and standardise the authoring and documenting of novel CHEM modules. A third module authoring library provides tools for disseminating themed bundles of CHEM modules as R libraries that are:

-	documented (with a website and PDF manuals);

-	licensed (using the copyleft GNU GPL-3 [66] by default);

-	easily citable (citation information can be retrieved within an R session or from hosting repositories); and

-	quality assured (each update triggers continuous integration workflows, including any acceptance and unit tests created by module library authors).

A library for *managing CHEM data* contains tools for supplying CHEM modules with input data ingested from local (i.e., a user's computer) or remote (online repositories) locations, labelling CHEM module datasets and exporting CHEM module data to online repositories. A library for *authoring reproducible analyses* contains tools to help write programs that apply CHEM modules to compatible datasets for the purpose of undertaking health economic analyses. These analysis programs can be either self-documenting (code is integrated with plain English explanations of what it does) or trigger the creation of explanatory documents (e.g., a scientific manuscript).

```{r, child = child_docs_ls$three, echo=FALSE}
```

##	Online services 
Framework libraries are designed to be used in conjunction with a number of third-party online services that we established and configured accounts with.

We created a GitHub organisation (a collection of code repositories) where source code that we author is stored and version controlled. We configured individual repositories in our GitHub organisation to implement continuous integration. By default, code libraries authored with our framework will use continuous integration to assess compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022]. To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022]. To facilitate the creation and hosting of documentation websites, we enabled GitHub Pages in each repository used for code library development.

We also created a Zenodo community - a collection of permanent, uniquely identified repositories. We then linked our Zenodo community and GitHub organisation so that every time we specify a version of code in one of our GitHub repositories as a “release”, a copy of that code is automatically created on Zenodo with a DOI. Finally, to manage model datasets, we created a dedicated collection within the Harvard Dataverse installation.

##	Documentation website
We developed a framework documentation website (https://www.ready4-dev.com) that provides guidance to model developers on how to use and contribute improvements to the ready4 software framework and CHEM modules developed with it. The documentation website is versioned, which means documentation relating to prior versions of framework software can be archived, retrieved and viewed.

The documentation website was developed using the Hugo framework [@hugo_2023], Docsy theme [@docsy_2023] and Algolia DocSearch [@algoliadocsearch_2023] and is hosted using the Netlify [@netlify_2023] service.  We used functions from our foundation framework library to partially automate website updates relating to available CHEM modules, datasets and analysis programs. We linked our Netlify account to our GitHub organisation so that the project website would automatically update whenever its  source code (publicly available in a GitHub repository) was edited.