### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd

---
title: "MS_Intro_Short"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four]. In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and software (a computational model).  Health economic scientific manuscripts typically describe a model using its conceptual and mathematical representations but report results that have been generated by its computational representation (i.e., the format that allows computers to perform calculations). The conceptual, mathematical and computational representations of a model are assumed to be isomorphic. However, the ease of independently assessing the validity of this assumption depends in part on how a health economic model has been implemented computationally.

The set of software files implementing a health economic model's structure, data and algorithms can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling software or open-source programming languages such as Python [@python2009] or R [@RCORE2022]. An advantage of commercial modelling software is that users often require limited or no programming skills to develop and apply CHEMs. However, open-source software programming languages may facilitate the development of CHEMs that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. The decision about what type of software development platform to use must balance ethical considerations and resource constraints (see Table \@ref(tab:proscons)).

```{r, child = child_docs_ls$one, echo=FALSE}
```

Open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]. The software used to develop CHEMs was identified as a barrier to open-source model implementations by participants in a 2020 survey of health economists [@Pouwels2022]. Other reported barriers relate to difficulties in generalising models, updating models and transferring model data, concerns about the level of public access and lack of interest from decision-makers, and legal, confidentiality and security considerations [@Pouwels2022].

<!-- Better software frameworks could potentially help address some of the barriers to open-source CHEM development. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources (e.g., code and data files), that can be modified to suit specific needs. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python.  -->

<!-- A high-level software framework for implementing open-source CHEMs in R has previously been developed with the primary aim of improving model transparency [@alarid2019need]. However, to gain widespread adoption, a software framework for CHEMs may need to provide tools to help model developers meet a broader range of ethical practice requirements. There are multiple challenges to overcome. Firstly, the ethical responsibilities of public health computational modellers remain poorly understood [@10.3389/fpubh.2017.00068] and the computational modelling landscape is changing fast [@calder2018computational]. Secondly, software frameworks can be challenging and time consuming to create and may become excessively complex over time [@edwin2014software]. Thirdly, software frameworks can be difficult to learn, often requiring model developers to undergo specialist training [@myllarniemi2018development]. -->