### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Tables.Rmd

---
title: "MA_Results"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(one = "../Child_RMDs/MS_Tables_1.Rmd",
                        two = "../Child_RMDs/MS_Tables_2.Rmd"
                        # ,
                        # three = "../Child_RMDs/MS_Tables_3.Rmd",
                        # four = "../Child_RMDs/MS_Tables_4.Rmd"
                        )
}else{
    child_docs_ls <- list(one = NULL,
                          two = NULL,
                          three = NULL,
                          four = NULL)
}
```