### 2024-02-14 Markdown/Manuscript/Child_RMDs/MS_Tables_2.Rmd

---
title: "Predr_Cors"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
```{r timelygls, tab.cap='Transparent, reusable and updatable (TRU) computational health economic models (CHEMs) are suggestive of ethical modelling practice', tab.id = 'timelygls', echo = F, results = 'asis', warning=FALSE}
library(kableExtra)  
df <- data.frame(` ` = c("Transparent", "Reusable", "Updatable"),
                 `Social acceptability` = c("Modeller judgments, model features and verification checks can be reviewed by third parties.",
                                                                        "Models can be modified to reflect alternative value judgments.",
                                                                        "Models can be modified to reflect alternative value judgments."),
                 `Fitness for purpose` = c("Modeller judgments, model features and verification checks can be reviewed by third parties.","Third party use increases likelihood of errors being identified.","Models can be validly used for longer."),
                 `Beneficial use` = c("Decision-makers can understand the strengths and weakness of models before applying them.","Can inform more decision problems, in more contexts with less duplicative effort.","Models can be validly used for longer."),
                 check.names = F)
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    # library(huxtable)
    # library(dplyr)
    #flextable::set_flextable_defaults() #fonts_ignore=TRUE
    # flextable::flextable(df) %>%
    #   flextable::border_inner(border = officer::fp_border(style = "none")) %>%
    #   flextable::border_outer() %>%
    #   flextable::align(align = "center", part = "header") %>%
    #   flextable::merge_h() %>%
    #   flextable::merge_v() %>%
    #   flextable::valign(j=2, valign = "top") %>%
    #   flextable::align(j=2:4, align = "center") %>%
    #   flextable::bold(j=1, part = "all") %>%
    #   flextable::bold(part = "header") %>%
    #   flextable::width(j = 1, width = 1.2) %>%
    #   flextable::width(j = 2:4, width = 1.6)
    # %>%
    #   
      #flextable::merge_at(i=2,j=1:2)
    # huxtable::as_hux(df) %>% 
    #   #insert_row("MPG-DISP", "", "", "HP-WT", fill = "") %>% 
    #   #huxtable::merge_cells(2:3,1) %>% 
    #  # huxtable::set_header_rows(1, TRUE) %>% 
    #   huxtable::set_col_width(c(.1,.1,.1,.7)) %>%
    #   huxtable::set_wrap(1, value=T)
      #huxtable::style_headers(bold = TRUE)
    df %>%
      kableExtra::kbl(booktabs = T,
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      kableExtra::column_spec(1, bold=T) %>%
      kableExtra::add_header_above(c(" ", "Project team responsibility" = 3), bold = T, italic = T) %>%
      kableExtra::column_spec(2, width = "13em") %>%
      kableExtra::column_spec(3, width = "13em") %>%
      kableExtra::column_spec(4, width = "13em") %>%
      kableExtra::collapse_rows(columns = 2, latex_hline = "none")
      # kableExtra::column_spec(3, width = "35em") %>%
      # kableExtra::pack_rows("Accountable - it is easy to see who developed, tested and applied a model and how they did it.",1,7) %>%
      # kableExtra::pack_rows("Updatable - a model and its components are maintained and continuously improved.",17,20) %>%
      # # kableExtra::pack_rows("Modular - models and their components can be combined to extend their scope",11,12) %>%
      # # kableExtra::pack_rows("Epitomising - model code can be re-used in other decision contexts",13,15) %>%
      # kableExtra::pack_rows("Reusable - a model and its components can be used in other models and by other modellers.",8,16) #%>%
      # kableExtra::pack_rows("Yielding - a model can be simply, flexibly and reliably used as a decision aid",19,20)
    }
```