### 2024-02-14 Markdown/Manuscript/Child_RMDs/MS_Tables_3.Rmd

---
title: "Predr_Cors"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
```{r cpkgs, tab.cap='Software framework R libraries', tab.id = 'cpkgs', echo = F, results = 'asis'}
library(kableExtra)  
df <- tibble::tibble(Package = c("ready4",
                                       "ready4fun",
                                       "ready4class",
                                       "ready4pack",
                                       "ready4use",
                                       "ready4show"),
                     Purpose = c("Provides a template model module and novel syntax and tools for finding interoperable CHEM modules, datasets and reproducible analysis programs.",
                                 "Streamline and standardise the authoring and documenting of functions that support selective reuse of model algorithms.",
                                 "Streamline and standardise the authoring and documenting of new CHEM modules.",
                                 "Help bundle and disseminate CHEM modules as documented, licensed and quality assured  R libraries.",
                                 "Help manage the labelling and transfer of data between CHEM modules and local and remote data repositories.",
                                 "Support the authoring of programs that implement integrated and reproducible data ingest, analysis and reporting pipelines."),
                           # A = c("3","2-3","2","1-2,5,7","1-2,7","4,6"), 
                           # R = c("4","5","4-5","1,3,6","3,7",""),
                           # U = c("","","","1-3","1",""),
                           `R library dependencies` = Package %>%
                             purrr::map(~ {
                               utils::packageDescription(.x) %>%
                                 `[`(c("Depends", "Imports")) %>%
                                 purrr::map(~{
                                   if(is.null(.x)){
                                     character(0)
                                     }else{
                                       .x %>%
                                         strsplit(",\\n") %>%
                                         purrr::flatten_chr() %>%
                                         purrr::map(~strsplit(.x,", ") %>%
                                                      purrr::flatten_chr()) %>%
                                         purrr::flatten_chr() %>% sort() %>%
                                         purrr::discard(~startsWith(.x,"R "))
                                       }
                                   }) %>%
                                 purrr::flatten_chr() %>%
                                 unique() %>%
                                 sort() %>%
                                 purrr::map_chr(~{
                                   updated_1L_chr <- stringr::str_replace_all(.x,"\\n"," ")
                                   problem_idx_1L_chr <- stringr::str_locate(updated_1L_chr," ")[1,1] %>%
                                     unname()
                                   if(!is.na(problem_idx_1L_chr))
                                     updated_1L_chr <- updated_1L_chr %>%
                                     stringr::str_sub(end = problem_idx_1L_chr-1)
                                   updated_1L_chr %>% 
                                     trimws(which = "left") 
                                   })
                               }) %>% 
                             purrr::map_chr (~ .x %>% paste0(collapse = " "))) %>% 
  as.data.frame() 
# if(params$X@outp_formats_chr[1] != "Word"){
#       df$Ref <- c(rep("",6)
#        # "@ready42022","@ready4fun2022","@ready4class2022","@ready4pack2022","@ready4use2022","@ready4show2022"
#                   )
# }else{
#   df$Ref <- ""
# }

df <- df %>%
  # dplyr::mutate(Standard = purrr::pmap_chr(df,
  #                                          ~paste0("A",
  #                                                 ..3,
  #                                                 paste0(rep("&nbsp;",
  #                                                     2+max(nchar(df[,3]))-nchar(..3)),
  #                                                     collapse = ""#
  #                                                     ),
  #                                                 ifelse(..4=="","","R"),
  #                                                 ..4,
  #                                                 paste0(rep("&nbsp;",
  #                                                     2+max(nchar(df[,4]))-nchar(..4)),
  #                                                     collapse = ""#
  #                                                     ),
  #                                                 ifelse(..5=="","","U"),
  #                                                 ..5,
  #                                                 paste0(rep("&nbsp;",
  #                                                     1+max(nchar(df[,5]))-nchar(..5)),
  #                                                     collapse = ""#
  #                                                     )#,
  #                                                 ))) %>%

  dplyr::select(Package,
                #Ref,
                Purpose,
                #Standard,
                `R library dependencies`)

names_chr <- names(df)
# df <- data.frame(df[,1],
#                  rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; &nbsp; &nbsp;"),
#                      nrow(df)),
#                  df[,2#:3
#                     ],
#                  rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; &nbsp; &nbsp;")
#                      ,nrow(df)),
#                  #df[,4],
#                  #rep("&nbsp; &nbsp; &nbsp;",nrow(df)),
#                  df[,3#5
#                     ])
# colnames(df) <- c("Package",
#                   ifelse(params$X@outp_formats_chr[1] == "Word","XXX"," "),
#                   #"Ref", 
#                   "Purpose", 
#                   ifelse(params$X@outp_formats_chr[1] == "Word","YYY"," "),
#                   #"Standard     ",# " ",
#                   "R library dependencies")
#df$Package <- "![](../Data/images/ready4.png){ width=10% }"#"\\includegraphics[valign=T,scale=2.5,raise=2mm]{../Data/images/ready4.png}"
#df$`Standard     `  <- df$`Standard     ` %>% stringr::str_replace_all("&nbsp;"," ")
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df %>% #[,c(1,3:4,6:7)] %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    #df$Ref<-NULL
    #df$Package <- ""
    # df$Package  <- df$Package %>% 
    #                   purrr::map_chr(~paste0("\\includegraphics[valign=M,width=1.9cm,raise=4mm]{../Data/images/",.x,"_logo.png}"))#scale=0.4,raise=3mm]
    # df$Package <- "\\includegraphics[valign=M,scale=0.08,raise=3mm]{../Data/images/ready4.png}"
    #df[,c(1,3,5:6)]   %>%
    df %>%
    # kableExtra::kbl(format = "markdown",
    #                   escape = F,
    #                 caption = knitr::opts_current$get("tab.cap"))
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      kableExtra::column_spec(2, width = "20em", latex_valign = "m") %>%
      # kableExtra::column_spec(1, 
      #                         image = kableExtra::spec_image(rep("../Data/images/ready4.png",nrow(df)), 
      #                                                           150, 150)) %>%
      kableExtra::column_spec(3, width = "20em",
                              latex_valign = "m")
    }
```