### 2024-02-15 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd

---
title: "MS_Discussion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

Ethical practice is a core expectation of health researchers and computational methods underpin most quantitative research. However, an understanding of the ethical dimensions about how health economic models are implemented as software is underdeveloped. We have attempted to address this gap by proposing that CHEM developers and funders adopt the ethical goals of socially acceptable user requirements and design specifications, fit for purpose implementations and socially beneficial post-release use. We suggest that CHEM attributes of transparency, reusability and updatability are indicative of projects that meet these goals and have identified six criteria for assessing these attributes.

Prior literature has noted that health economic models routinely have limitations in transparency [@Jalali2021; @McManus2019; @Bermejo2017; @Ghabri2019], reuse [@Feenstra2022; @Emerson2019], and being kept up to date [@Sampson2017; @10.3389/fpubh.2022.899874]. These limitations not only pose ethical risks such as those we have described, but are practical barriers to model generalisability and transferability. Addressing these limitations will require changes to the work practices, funding and tools that underpin health economic model development.

Adherence to good practice guidance is an essential requirement for healthcare modelling [@Erdemir2020] and there is a need for better and comprehensive guidelines about how health economists should implement health economic models as software. Research funders need to actively support initiatives to help remove barriers to better practice in this area. A core responsibility of research funders is reducing waste in research [@chalmers2014increase]. Investments in modelling projects might generate greater social returns if they supported the project activities necessary to develop CHEMs that are easily understood, widely reused and well maintained.

Previously recommended strategies for more beneficial health economic research investments include support for harmonised ethical standards for model development [@10.3389/fpubh.2017.00068], methodological innovation to improve model transferability [@craig2018taking], networks of modellers working on common health conditions [@Sampson2019], and  centralized infrastructure such as open source model repositories [@Pouwels2022] and a standard platform for model implementations [@Ghabri2019]. Each of these strategies could benefit from enabling technical infrastructure, for example software frameworks that support the development of CHEMs that meet ethical criteria. We are currently engaged in developing a prototype of one such software framework [@ready4fw2024].

<!-- Development of software frameworks to support ethical CHEM implementations could enable and enhance each of these strategies. -->

<!-- As illustrated by Table \@ref(tab:checktb), we have developed a prototype software framework that can be used to implement youth mental health CHEMs that largely satisfy our TRU criteria. However, in its current form our software framework is too fragile to be recommended as anything more than a prototype for supporting the development needs of other modelling teams and projects. We developed our prototype software framework primarily to address the user requirements of one group – ourselves. Factors such as user enjoyment, usability, active user-community and supporting resources influence adoption of software frameworks [@myllarniemi2018development]. We currently lack the resources required to  address these issues for our software framework. -->

<!-- However, our prototype has a number of features that future work on CHEM software frameworks may find useful to incorporate. Firstly, developing a software framework to work within an existing and widely used open source programming language such as R, can keep framework scope relatively narrow. This makes it more tractable to develop, maintain and learn, and facilitates integration with other modelling tools written in that language (e.g., the dependency libraries we list in Table \@ref(tab:cpkgs)). Secondly, implementation that combines both object-oriented and functional programming paradigms can avail of the modular and syntactical simplicity benefits of the former, while limiting needless bundling of code artefacts. Thirdly, a sensible trade-off needs to be found between transparent code implementation (which requires clear and sufficiently detailed documentation) and Agile Software Development (for which a foundational principle is prioritising the development of working code over writing documentation [@beck2001manifesto]). Our software framework makes this trade off by enforcing the use of consistent code style conventions and file organisation which in turn enables automated generation of simple documentation at every CHEM module update. All model data-structures and algorithms are therefore always documented (at least minimally, with machine authored content), meaning model developers have a requirement to write customised documentation less frequently. Many of the features of our prototype software framework have the potential to be significantly enhanced through integration with evolving code and documentation authoring capabilities of large language models. -->

<!-- Our software framework provides the outline of a technical solution to some of the challenges of open source CHEM development. However, better tools alone are insufficient and policy and institutional changes are also required.  -->

<!-- A future software framework for ethical CHEMs would ideally incorporate a base set of features useful to developers of computational models across all domains of public health and health economics, with the capability for community-led extensions that are tailored to the needs of modellers focused on specific health-conditions. Such approaches will likely be important to ensure a sufficient potential user-base to make CHEM software frameworks sustainable. -->