### 2024-02-15 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd

---
title: "MS_Imp_Guidelines"
output: html_document
date: '2022-07-21'
---

<!-- We considered literature on modelling practice and reflected on our prior experience with developing and using CHEMs to identify: (i) some core ethical responsibilities of CHEM project teams; (ii) attributes of CHEMs that can suggest fulfillment of these responsibilities; and (iii) criteria against which these attributes can be assessed.  -->
<!-- responsibilities to goals -->
# Ethical goals for CHEM developers and funders
Boden and McKendrick [@10.3389/fpubh.2017.00068] propose a framework for ethical public health modelling based on the criteria of independence (concerning how modeller subjectivity shapes model design), beneficence / non-maleficence (concerning model quality and utility), transparency (concerning the need for policymakers to reliably evaluate model strengths and weaknesses) and justice (concerning the social obligations of modellers to consider and communicate ethical issues about model use). <!-- However, t -->These criteria are not specific to the computational model, but <!-- are designed to  -->apply to a modelling project as a whole, including a model's conceptual and mathematical representations. The authors propose 13 questions to evaluate ethical risk across the four criteria, but of these, only "is the model code open source or available on request" (for transparency) and "has the model been verified, i.e., does it do what the modeller wants it to do?" (for beneficence / non-maleficence) are specific to model's representation as software.

To identify ethical issues that are specific to CHEMs, we <!-- first  -->considered <!-- Edit -->how Boden and McKendrick's four criteria might apply to <!-- /Edit -->the distinct phases in a software development project. Software development lifecycle (SDLC) models take different approaches to the inclusion, naming, definition, sequencing and iteration of project phases [@ruparelia2010software]. However, SDLC models typically have components that can principally map to the concepts of *planning* (e.g., identification of system / user requirements and design specifications), *implementation* (e.g., development and testing) and *release* (e.g., deployment, system integration, maintenance and support). For simplicity, we propose one overarching ethical <!-- responsibility  for CHEM project team-->goal for each phase. These goals<!-- responsibilities  --> are: 
<!-- establish  -->
(i) *socially acceptable user requirements and design specifications* (during CHEM planning);
<!-- ensure  -->
(ii) *fitness for purpose* (of CHEM implementations); and
<!-- support  -->
(iii) *socially beneficial use* (post CHEM release).

Boden and McKendrick’s transparency criterion underpins the ethical goals we propose for all three phases (Table \@ref(tab:timelygls)). Boden and McKendrick’s other <!-- ethical  -->criteria can be selectively mapped to specific phases: our proposed goal for ethical CHEM planning incorporates justice and independence considerations; the goal we propose for CHEM implementation is based on independence and beneficence / non-maleficence criteria; and our proposed ethical goal post CHEM release reflects beneficence / non-maleficence and justice criteria. <!-- /Edit -->

Misalignment between the values of computational model developers and those of the population groups affected by decisions based on their models presents significant ethical risks [@thompson2022escape; @duckett2022journey].  The value judgments of health economic modellers are rarely adequately specified [@duckett2022journey]. These value judgments influence the assumptions, selection of model features and standards for evidence that shape health economic model projects [@HARVARD2020112975]. For example, the process of determining CHEM *user requirements and design specifications* will involve making judgments about the competing claims of alternative software development platforms (Table \@ref(tab:proscons)). That decision will in turn determine how straightforward it is for a CHEM to be subsequently amended to reflect alternative value judgments relating to model conceptualisation and mathematical formalisation (Table \@ref(tab:timelygls)).
<!-- Moved -->
<!--  , omissions that may lead to socially unacceptable policy recommendations -->

Modellers need to do more than self-certify that their <!-- have duties both to ensure a  -->computational model is *fit for purpose* - <!-- Edit -->they also need to<!-- /Edit --> provide potential third-party users with the means of assessing its adequacy 
<!-- for their proposed uses  -->[@Erdemir2020; @Feenstra2022; @thompson2019escape; @10.3389/fpubh.2017.00068]. However, it is common for health economic models to have serious methodological flaws [@carletto_zanuzzi_sammarco_russo_2020; @WONDER2015467]; insufficient validation [@Ghabri2019; @kolovos2017model; @haji2013model], poor reproducibility [@Jalali2021; @McManus2019; @Bermejo2017]; and undeclared errors [@Radeva2020]. Appropriate computational implementation choices can help address many of these shortcomings, for example by automating quality assurance checks and <!-- Edit -->providing executable code for  <!--facilitating manual  -->review and testing <!-- /Edit -->by third parties.

The *social benefit* of developing a CHEM may be limited if it is rarely used, if mis-used or when its acceptability and adequacy rapidly decay. Reuse of CHEMs as components of other models can potentially make model development more efficient [@Arnold2010; @garcia2021cost]. However, health economic models face challenges related to transferability across jurisdictions [@garcia2021cost] that create barriers to reuse. Without ongoing maintenance, a CHEM risks becoming less reliable with time [@garcia2021cost] and is at risk of being deployed for purposes for which it is poorly suited [@calder2018computational]. Currently, health economic models are rarely implemented computationally in a manner that facilitates routine updates [@Sampson2017], thus limiting the temporal window within which a CHEM can be validly applied. 
<!-- Moved -->
<!-- Edit --><!-- All three  -->Our proposed ethical goals are most likely to be met when there is both good<!-- have implications for both  --> professional practice by CHEM developers and <!-- Addition -->adequate <!-- Addition -->resourcing by project funders.  <!-- /Edit --> The choices CHEM developers make often need to balance ethical considerations with project resource constraints (as illustrated in Table \@ref(tab:proscons)). Modelling projects implemented by diligent and conscientious CHEM developers may still be exposed to ethical risks if there are inadequate resources to meaningfully engage important stakeholder groups, prepare code and data artefacts for independent assessment and provide ongoing model maintenance and user support. Existing incentive structures for research health economists generally do not promote facilitating peers to reuse their work and funding is rarely adequate to maintain and routinely update a model. Currently, it can take “an extraordinary amount of idealism” to commit to authoring and maintaining research software [@anzt2020environment]. For these reasons, we believe responsibility for achieving our proposed ethical goals is shared between CHEM developers and funders. Reflecting the shared nature of this responsibility, we suggest that in most cases where a modelling project fails to meet our ethical goals, it will be more appropriate to describe these shortcomings as "ethical risks" (the language used by Boden and McKendrick) than to use accusatory terms like "unethical practice."<!-- these responsibilities should apply jointly to health economic model developers and funders. -->

<!-- For example, as illustrated in Table \@ref(tab:proscons), project resource constraints may require model developers to make trade-offs -->
<!-- As it is unreasonable to place expectations of modellers that they are not resourced to fulfil,  -->
<!-- /Moved -->

<!-- # -->
# CHEM attributes that can reduce ethical risk
<!-- The responsibilities of model project teams over a CHEM lifecycle  -->
Our proposed ethical goals for each project phase are easier to state than to measure.  However, <!-- aspects of ethical modelling practice  -->achievement of these goals may be inferred from measurable attributes of CHEMs. As described in Table \@ref(tab:timelygls),  we believe that the <!-- Edit -->developers and funders<!-- /Edit --> of CHEMs that are transparent, reusable and updatable (TRU) are more likely to have achieved the ethical goals we propose. <!-- fulfilled a number of their ethical obligations.   --> 
<!-- We therefore selected these CHEM attributes to use as the basis for deriving ethical assessment criteria. -->

```{r, child = child_docs_ls$two, echo=FALSE}
```

Taking steps to make computational models *transparent* is an important means of reducing ethical risk <!-- part of ethical public health modelling practice  -->[@10.3389/fpubh.2017.00068]. Guidance on transparency in health economic modelling recommend that model code and data should be clearly documented [@Eddy2012]. Notably, the same guidelines, published over ten years ago, do not recommend sharing model code and data. However, more recent healthcare modelling guidance does recommend public dissemination of such artefacts [@Erdemir2020]. Online repository services such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. Ensuring that calculations are correct and consistent with model specifications is an essential part of CHEM quality assurance [@techver2019]. The extensiveness of such verification checks can be reported using the concept of code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. Tests should ideally combine both unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). The nature and extent of individual model authorship contributions can become unclear when models are implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape]. This issue can be addressed by use of online repository services such as GitHub [@github2007], that provide citation tools and can transparently record all individual code contributions to a modelling project over its lifecycle.

<!-- A CHEM that is *reusable* also signals ethical modelling practice.  -->
Although making a CHEM’s code, data and documentation publicly available is increasingly considered good practice, it is insufficient to ensure that a model is *reusable*. Key concerns for health economists when considering whether to reuse a model are generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is good scientific computing practice [@Wilson_2017] and promotes selective reuse. Computational implementations that store model code and data in distinct files and locations (as opposed to embedding data such as parameter values into source code) are also easier to selectively modify. Modular implementations enable models to be constructed from multiple independently reusable and replaceable sub-models (modules) [@pan2021modular]. The programming concept inheritance [@8717448], deployed in object-oriented programming approaches, can facilitate the selective reuse and modification of model components. Granting permissions to others to use, test and adapt models and their components, can be facilitated by two broad categories of open source licenses. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. Alternatively, copyleft licenses [@copyleft2022] can be used to require that derivative works created by content users remain open-source.

Models should be *updatable* so that they remain valid for longer, evolving as new evidence emerges and the systems being modelled change [@Jenkins2021; @garcia2021cost]. Ensuring that a model is regularly reviewed to identify and implement required improvements is a recommended defence against model validity decay [@calder2018computational]. Sustainable maintenance of open source research software requires both a core development team and an active user community [@info:doi/10.2196/20028]. Online communities can be an efficient means of engaging model users in testing each version of a model, identifying issues and suggesting improvements. Services such as GitHub [@github2007] provide collaborative code development tools [@MERGEL2015464] that help elicit, integrate and reconcile contributions from multiple contributors and to ensure each update is uniquely identifiable and retrievable. It is also important that verification checks are rerun with each model update, a task that can be automated using the software development practice of continuous integration [@CI2017].  The risk of software revisions having unintended consequences for third party users can be mitigated through the use of deprecation conventions [@zhou2016api] that take an informative and staged approach to retiring outdated CHEMs.

<!-- # -->
#	Criteria for assessing TRU attributes
We suggest six criteria (two for each attribute) against which for assessing the TRU attributes of CHEMs.

How *transparent* a CHEM is can be assessed against the criteria:

-	*T1*: Software files (executable code, testing procedures and outcomes and non-confidential data) are available in open access repositories.

-	*T2*: It is easy to see who funded, developed and tested each part of the CHEM and to identify the modelling team’s assumptions, judgments and theories about CHEM development and use.

How *reusable* a CHEM is can be assessed against the criteria:

-	*R1*: Programming practices facilitate selective reuse of model sub-components.<!-- Replacement -->

-	*R2*: Licenses restrict only unethical use and allow anyone to ethically reuse model code and non-confidential data, in whole or in part, without charge, and for purposes that include the creation of derivative works.

How *updatable* a CHEM is can be assessed against the criteria:

-	*U1*: Maintenance infrastructure is in place to support version control and collaboration with model users.

-	*U2*: Each new release of a CHEM is systematically retested, with changes implemented to minimise disruptions for existing model users.