### 2024-03-19 Markdown/Manuscript/Child_RMDs/MS_Cndl_Tabs_Not_In_Body.Rmd

---
title: "MS_Cndl_Child_Not_In_Body"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

\newpage
<!-- `r if(!is.null(child_tabs_1L_chr)){"## Tables {-}"}` -->

<!-- and figures -->

```{r, child = '../Child_RMDs/MS_Tables_1.Rmd', echo=FALSE}
```

\newpage

```{r, child = '../Child_RMDs/MS_Tables_2.Rmd', echo=FALSE}
```

\newpage

```{r, child = '../Child_RMDs/MS_Tables_3.Rmd', echo=FALSE}
```
\newpage
<!-- # ```{r, child = '../Child_RMDs/MS_Tables_4.Rmd', echo=FALSE} -->
<!-- # ``` -->