### 2024-03-19 Markdown/Manuscript/Child_RMDs/MS_Conclusion.Rmd

---
title: "MS_Conclusion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
`r ifelse(is.na(params$X@conclusion_1L_chr),"",params$X@conclusion_1L_chr)` 
<!-- We have identified criteria that can be used to systematically assess the extent to which the computational implementation of health economic models adheres to transparency, reusability and updatability criteria. <!-- We have developed a prototype software framework that can support the implementation of CHEMs that largely meet these criteria. Our framework can be used as a prototype for developing future software frameworks to support ethical implementation of CHEMs.  --> 
<!-- Barriers to ethical CHEM practice arising from resource constraints should be addressed by research funders. -->
 We propose that CHEM developers and funders adopt the ethical goals of socially acceptable user requirements and design specifications, fit for purpose implementations and socially beneficial post-release use. We suggest that CHEM attributes of transparency, reusability and updatability are indicative of projects that meet these goals and have identified six criteria for assessing these attributes. We hope our proposals contribute to a timely and much needed debate within the health economics community about the ethical dimensions of implementing health economic models as software.