### 2024-03-20 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd

---
title: "MS_Intro_Short"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four]. In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and software (a computational model). Health economic scientific manuscripts typically involve description of a model using its conceptual and mathematical representations and report results that have been generated by its computational representation (i.e., the format that allows computers to perform calculations). Authors of these manuscripts assume (rarely explicitly) that the conceptual, mathematical and computational representations of a model are internally consistent and externally valid. The feasibility of independently assessing the validity of this assumption depends in part on how a health economic model has been implemented computationally.

The set of software tools and files implementing a health economic model's structure, data and algorithms can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling <!-- Addition -->or spreadsheet<!-- /Addition --> software (for example, Microsoft Excel [@msexcel]) or open-source programming languages such as Python [@python2009] or R [@RCORE2022]. Training and professional practice guidance for health economists typically provide only superficially coverage of software development topics. However, these capabilities will be critical for achieving proposed methodological innovations at the frontier of the health economics field. Opportunities for field development that closely overlap with CHEM implementation approaches include living health technology assessments [@thokala2023living], machine learning powered health economic analyses [@padula2022machine], open source health-economic models [@Pouwels2022] and utilising Large Language Models (LLMs) to program health economic models [@poirrier2023msr26]. <!-- Potential enhancements to health economic modelling practice include improved model transferability [@munthe2019systematic; @barbieri2010international], greater use of reference models [@Afzali2013], living health technology assessments [@thokala2023living] and leveraging the analytic power of machine learning [@padula2022machine]. -->

There is a significant ethical dimension to software development. For example, recent developments in LLMs mean software developers may confront novel ethical issues including: new security risks; more restrictive open-source licensing; environmentally harmful levels of computing power; and concerns about model explainability and unintended consequences [@ozkaya2023application]. The health economics community should engage more directly with these ethical debates.  The computational modelling landscape is changing fast [@calder2018computational] and the ethical <!-- responsibilities  --><!-- Edit -->dimensions of computational public health modelling <!-- /Edit -->remain poorly understood [@10.3389/fpubh.2017.00068]. Documents such as the World Health Organisation's recent guidance on generative artificial intelligence [@world2024ethics] can be important reference points in helping the health economics discipline navigate relevant ethical topics in software development.

  <!-- Moved -->
 <!-- However, open-source CHEMs remain rare [@Feenstra2022; @Emerson2019], despite in principle support from many in the health economics field [@Pouwels2022]. -->
 <!-- Moved -->

<!-- Health economists must balance ethical considerations and resource constraints when deciding what type of software development platform to use for authoring CHEMs. <!-- ( see--><!-- Addition --> 
<!--In Table \@ref(tab:proscons), we summarise some of the trade-offs we consider when undertaking modelling projects<!-- ) --> <!-- Edit Move --><!--Determining how to make these trade-offs can be difficult.  -->

 <!-- /Edit Move-->

```{r, child = child_docs_ls$one, echo=FALSE}
```

<!--  The software used to develop CHEMs was identified as a barrier to open-source model implementations by participants in a 2020 survey of health economists [@Pouwels2022]. Other reported barriers relate to difficulties in generalising models, updating models and transferring model data, concerns about the level of public access and lack of interest from decision-makers, and legal, confidentiality and security considerations [@Pouwels2022].  -->

<!-- Better software frameworks could potentially help address some of the barriers to open-source CHEM development. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources (e.g., code and data files), that can be modified to suit specific needs. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python.  -->

<!-- A high-level software framework for implementing open-source CHEMs in R has previously been developed with the primary aim of improving model transparency [@alarid2019need]. However, to gain widespread adoption, a software framework for CHEMs may need to provide tools to help model developers meet a broader range of ethical practice requirements. There are multiple challenges to overcome. -->

<!-- Secondly, software frameworks can be challenging and time consuming to create and may become excessively complex over time [@edwin2014software]. Thirdly, software frameworks can be difficult to learn, often requiring model developers to undergo specialist training [@myllarniemi2018development]. -->