### 2024-03-20 Markdown/Manuscript/Child_RMDs/MS_Tables_2.Rmd

---
title: "Predr_Cors"
author: "Author"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r echo=FALSE}
df <- data.frame(` ` = c("Transparent", "Reusable", "Updatable"),
                 `Social acceptability` = c("Modeller judgments, model features and verification checks can be reviewed by third parties.",
                                                                        "Models can be modified to reflect alternative value judgments.",
                                                                        ""#"Models can be modified to reflect alternative value judgments."
                                            ),
                 `Fitness for purpose` = c("",#"Modeller judgments, model features and verification checks can be reviewed by third parties.",
                                           "Third party use increases likelihood of errors being identified.",
                                           "Models can be validly used for longer."),
                 `Beneficial use` = c("Decision-makers can understand the strengths and weakness of models before applying them.",
                                      "Can inform more decision problems, in more contexts with less duplicative effort.",
                                      ""#"Models can be validly used for longer."
                                      ),
                 check.names = F)
```
```{r echo=FALSE}
if(params$X@outp_formats_chr[1] == "Word"){
  conditional_1L_chr <- "../Child_RMDs/MS_Tables_2_WORD.Rmd"
}else{
    conditional_1L_chr <- "../Child_RMDs/MS_Tables_2_LATEX.Rmd"
}
```
```{r child = conditional_1L_chr, echo=FALSE}
```