### 2024-03-20 Markdown/Manuscript/Child_RMDs/MS_Tables_2_WORD.Rmd

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
date: "2024-03-20"
---

```{r timelygls, tab.cap='Transparent, reusable and updatable (TRU) computational health economic models (CHEMs) are suggestive of ethical modelling practice', tab.id = 'timelygls', echo = F, results = 'asis', warning=FALSE}
library(kableExtra)

 library("flextable")
  library("ftExtra")
if(params$X@outp_formats_chr[1] == "Word") {

  df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    #library(huxtable)
    # library(gt)
    # library(dplyr)
   # flextable::set_flextable_defaults() #fonts_ignore=TRUE
    # flextable::flextable(df) %>%
      # flextable::border_inner(border = officer::fp_border(style = "none")) %>%
      # flextable::border_outer() %>%
      # flextable::align(align = "center", part = "header") 
    # %>%
    #   flextable::merge_at(i = 3, j = 3:4)
      # flextable::merge_h() %>%
      # flextable::merge_v() %>%
      # flextable::valign(j=2, valign = "top") %>%
      # flextable::align(j=2:4, align = "center") %>%
      # flextable::bold(j=1, part = "all") %>%
      # flextable::bold(part = "header") %>%
      # flextable::width(j = 1, width = 1.2) %>%
      # flextable::width(j = 2:4, width = 1.6)
    # %>%
    #   
      #flextable::merge_at(i=2,j=1:2)
    # gt::gt(df) %>%
    #    gt::cols_width(
    # ` ` ~ gt::px(50),
    # `Social acceptability` ~ gt::px(120),
    # `Fitness for purpose` ~ gt::px(120),
    # `Beneficial use` ~ gt::px(120)#,
    #gt::everything() ~ gt::px(60)
       #) # %>%
    #   gt::cols_merge(columns = `Social acceptability`,
    #                  rows = 2:3) %>%
      # gt::cols_merge(columns = c(`Fitness for purpose`, `Beneficial use`),
      #                rows = 1)

    # huxtable::as_hux(df) %>%
    #   huxtable::to_latex(tabular_only = TRUE) 
    #%>%
      #insert_row("MPG-DISP", "", "", "HP-WT", fill = "") %>%
     #  huxtable::merge_cells(2:3,1) %>%
     # huxtable::set_header_rows(1, TRUE) %>%
      # huxtable::set_col_width(c(.1,.1,.1,.7)) %>%
      #huxtable::set_wrap(1, value=T) %>%
   # huxtable::style_headers(bold = TRUE)
    # df %>%
    #   kableExtra::kbl(booktabs = T,
    #                   escape = F,
    #                   caption = knitr::opts_current$get("tab.cap")) %>%
    #   kableExtra::kable_paper(full_width = F) %>%
    #   kableExtra::column_spec(1, bold=T) %>%
    #   #kableExtra::add_header_above(c(" ", "Project team responsibility" = 3), bold = T, italic = T) %>%
    #   kableExtra::column_spec(2, width = "13em") %>%
    #   kableExtra::column_spec(3, width = "13em") %>%
    #   kableExtra::column_spec(4, width = "13em") %>%
    #   kableExtra::collapse_rows(columns = 2, latex_hline = "none")

    }
```