### 2024-03-20 Markdown/Manuscript/Parent_PDF/Test_Tb.Rmd

---
output:
  bookdown::pdf_book:
    base_format: rticles::springer_article
    # includes:
    #   in_header: preamble.tex
    keep_tex: true
#     pandoc_args: 
#       - '--lua-filter=../Filters/scholarly-metadata.lua'
#       - '--lua-filter=../Filters/author-info-blocks.lua'
classoptions: 
  - sn-vancouver      
  - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
  # - referee       # Optional: Use double line spacing 
  # - lineno        # Optional: Add line numbers
  # - iicol         # Optional: Double column layour

title: Identifying and assessing ethical risks when implementing health economic models as software
titlerunning: Ethical computational modelling

authors: 
  - firstname: Matthew
    particle: P 
    lastname: Hamilton
    email: matthew.hamilton1@monash.edu
    affiliation: [1,2]
    corresponding: TRUE
  - firstname: Caroline
    lastname: Gao
    affiliation: [2,4,1]

  - firstname: Jonathan  
    lastname: Karnon
    affiliation: [3]

  - firstname: Luis  
    lastname: Salvador-Carulla
    affiliation: [5]

  - firstname: Sue
    particle: M
    lastname: Cotton
    affiliation: [2,4]

  - firstname: Cathrine
    lastname: Mihalopoulos
    affiliation: [1]

affiliations:
  - number: 1
    corresponding: TRUE
    info:
      orgdiv: School of Public Health and Preventive Medicine
      orgname: Monash University
    address:
        city: Melbourne
        country: Australia
  - number: 2
    corresponding: FALSE
    info:
      orgname: Orygen
    address:
        city: Parkville
        country: Australia
  - number: 3
    corresponding: FALSE
    info:
      orgname: Flinders University   
    address:
        city: Adelaide
        country: Australia 
  - number: 4
    corresponding: FALSE
    info:
      orgdiv: Centre for Youth Mental Health
      orgname: The University of Melbourne
    address:
        city: Parkville
        country: Australia

  - number: 5
    corresponding: FALSE
    info:
#      orgdiv: XXXXXXXXXXX
      orgname: University of Canberra  
    address:
    #     # street: "182 Victoria Square"     
        city: Canberra
    #     postcode: 5000
    #  #   state: South Australia
        country: Australia 
keywords:  
  - computational models
  - ethics of modelling
  - health economics
  - open-source models

pacs: 
  jel:
    - "C63"
    - "C88"
    - "I10"
  msc:
    - "91-08"
# title: "REPLACE ME"
params:
  X: null
abstract: TO BE REPLACED
# output: 
#   pdf_document:
#     keep_tex: true
graphics: true
header-includes: |
  %% Per the spinger doc, new theorem styles can be included using built in style, 
  %% but it seems the don't work so commented below
  %\theoremstyle{thmstyleone}%
  \newtheorem{theorem}{Theorem}%  meant for continuous numbers
  %%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
  %% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
  \newtheorem{proposition}[theorem]{Proposition}%
  %%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

  %% \theoremstyle{thmstyletwo}%
  \theoremstyle{remark}
  \newtheorem{example}{Example}%
  \newtheorem{remark}{Remark}%

  %% \theoremstyle{thmstylethree}%
  \theoremstyle{definition}
  \newtheorem{definition}{Definition}%

  \usepackage{float}
  \let\origfigure\figure
  \let\endorigfigure\endfigure
  \renewenvironment{figure}[1][2] {
      \expandafter\origfigure\expandafter[H]
  } {
      \endorigfigure
  }
  \usepackage{lscape}
  \usepackage{pdflscape}
  \newcommand{\blandscape}{\begin{landscape}}
  \newcommand{\elandscape}{\end{landscape}}
  \usepackage[none]{hyphenat}
  \usepackage[export]{adjustbox}
  \usepackage{multirow}
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{wrapfig}
  \usepackage{colortbl}
  \usepackage{booktabs} 
  \usepackage{colortbl}
  \usepackage{pdflscape}
  \usepackage{tabu}
  \usepackage{threeparttable}
  \usepackage[normalem]{ulem}
  \usepackage{threeparttablex}
  \usepackage{makecell}
  \expandafter\def\csname
  ver@l3backend.sty\endcsname{}
---

```{r echo = F}
## NOTE: EDITED TEMPLATE FILE To Add rotating and multirow packages as per:
# https://stackoverflow.com/questions/31503418/how-to-add-a-latex-package-to-r-journal-template-from-rticle-package-in-r
library(ready4show)
options(tinytex.verbose = TRUE)
# library(xtable)
# options(xtable.floating = FALSE)
# options(xtable.timestamp = "")
```
```{r}
params_ls <- list(X = ready4show::Ready4showSynopsis())
```

```{r timelygls, tab.cap='Transparent, reusable and updatable (TRU) computational health economic models (CHEMs) are suggestive of ethical modelling practice', tab.id = 'timelygls', echo = F, results = 'asis', warning=FALSE, message=FALSE}
library(kableExtra)  
df <- data.frame(` ` = c("Transparent", "Reusable", "Updatable", "Blank"),
                 `Social acceptability` = c("Modeller judgments, model features and verification checks can be reviewed by third parties.",
                                                                        "Models can be modified to reflect alternative value judgments.",
                                                                        "",#"Models can be modified to reflect alternative value judgments.",
                                            "XXXXXXXXXXXXXXXXX"
                                            ),
                 `Fitness for purpose` = c("",#"Modeller judgments, model features and verification checks can be reviewed by third parties.",
                                           "Third party use increases likelihood of errors being identified.",
                                           "Models can be validly used for longer.",
                                           "XXXXXXXXXXXXXXXXX"),
                 `Beneficial use` = c("Decision-makers can understand the strengths and weakness of models before applying them.",
                                      "Can inform more decision problems, in more contexts with less duplicative effort.",
                                      "",#"Models can be validly used for longer."
                                      "XXXXXXXXXXXXXXXXX"),
                 #Blank = c(rep("XXXXXXXXXXXXXXXXXXXXXXXXXXXX",3)),
                 check.names = F)
 library("flextable")
  library("ftExtra")
if(params_ls$X@outp_formats_chr[1] == "Word") {

  df %>%
    ready4show::print_table(output_type_1L_chr = params_ls$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params_ls$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    # library(huxtable)
    # library(gt)
    # library(dplyr)
   #flextable::set_flextable_defaults() #fonts_ignore=TRUE
   flextable::flextable(df) %>%
   flextable::border_inner(border = officer::fp_border(style = "none")) %>%
   flextable::border_outer() %>%
   flextable::align(align = "center", part = "header") %>%
   #flextable::padding(j=2:4, part = "all", padding.top = 40, padding.bottom = 40) %>%
   # flextable::merge_h() %>%
   # flextable::merge_v() %>%
   flextable::align(j=2:4, align = "center") %>%
   flextable::bold(j=1, part = "all") %>%
   flextable::bold(part = "header") %>%
   flextable::height(i = 1:3, height = 1.2) %>%
   flextable::hrule(rule = "exact", part = "body") %>% 
   flextable::width(j = 1, width = 1) %>%
   flextable::width(j = 2:4, width = 1.8) %>%
   flextable::merge_at(i=1,j=2:3) %>%
   flextable::merge_at(i=2:4,j=2) %>%
   flextable::merge_at(i=3:4,j=1) %>%
   flextable::merge_at(i=3:4,j=3:4)  
    }
```