### 2024-03-25 Markdown/Manuscript/Child_RMDs/MS_Tables_1.Rmd

---
title: "Predr_Cors"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
```{r proscons, tab.cap='Implications of choosing proprietary software or an open-source programming language for the computational implementation of health economic models', tab.id = 'proscons', echo = F, results = 'asis'}
library(kableExtra)  
df <- data.frame(`Implications` = c(rep("Resources",4), rep("Ethical",10)),
                 `Proprietary modelling or spreadsheet software` = c("(+) Will include tools to efficiently accomplish many common modelling tasks.",
                                                     "(+) Requires limited or no programming skills.",
                                                     "(-) Requires model developers and users to purchase (in some cases, expensive) software licenses.",
                                                     "",
                                                     "Transparency",
                                                     "(+) Popular tools (e.g. Excel) are readily understood by many.",
                                                     "(-) Manual transfer of results to manuscripts through copy and paste or transcription from visual inspection increases risk of (undetected) reporting errors.",
                                                     "",
                                                     "Reusability",
                                                     "(+) Files will reliably open and execute correctly for years after project is completed.",
                                                     "",
                                                     "",
                                                     "Updatability",
                                                     "(-) Modifications and adaptations, particularly by third parties, may require bespoke permissions and it can be difficult to clearly itemise all changes."),
                 `Open-source programming language` = c("(-) Initial development is likely to be time and resource intensive.",
                                                        "(+) Over medium term, efficiency savings are possible if:",
                                                        "(i) artefacts (code and data) from pre-existing models are leveraged; and / or",
                                                        "(ii)	models are to be maintained and transferred (e.g. to new decision problems / contexts).",
                                                        "Transparency",
                                                        "(+) Easy to publicly share testable source code while selectively restricting access to confidential data.",
                                                        "(+) Easy to make clear who has developed and tested each part.",
                                                        "(-) Novel code increases likelihood of (undetected) software errors.",
                                                        "Reusability",
                                                        "(-) Potentially fragile – if not maintained / not bundled with all required and correctly versioned dependencies.",
                                                        "(+) Facilitates modular implementations that make complex model representations more tractable.",
                                                        "(+) Facilitates customised integration with existing data / decision systems.",
                                                        "Updatability",
                                                        "(-) Facilitates collaborative development and maintenance."
                                                        ),
                 check.names = F)
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    df %>%
      dplyr::filter(!`Proprietary modelling or spreadsheet software` %in% c("Transparency", "Reusability","Updatability")) %>%
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      #kableExtra::pack_rows("Resource requirements implications",1,4, hline_after = T) %>%
      #kableExtra::pack_rows("Ethical implications",5,14, hline_after = T) %>%
      kableExtra::column_spec(1, width = "7em", bold=T) %>% #, latex_valign = "m"
      kableExtra::column_spec(2, width = "20em") %>%
      kableExtra::column_spec(3, width = "20em") %>%
      kableExtra::collapse_rows(columns = 1, latex_hline = "major") %>%
      kableExtra::pack_rows("Transparency",5,7, latex_align = "c") %>% #bold = F, italic = T
      kableExtra::pack_rows("Reusability",8,10, latex_align = "c") %>%
      kableExtra::pack_rows("Updatability",11,11, latex_align = "c") 
      # kableExtra::pack_rows("Accountable - it is easy to see who developed, tested and applied a model and how they did it.",1,7) %>%
      # kableExtra::pack_rows("Updatable - a model and its components are maintained and continuously improved.",17,20) %>%
      # kableExtra::pack_rows("Modular - models and their components can be combined to extend their scope",11,12) %>%
      # kableExtra::pack_rows("Epitomising - model code can be re-used in other decision contexts",13,15) %>%
      # kableExtra::pack_rows("Reusable - a model and its components can be used in other models and by other modellers.",8,16) #%>%
      # kableExtra::pack_rows("Yielding - a model can be simply, flexibly and reliably used as a decision aid",19,20)

    }
```